---
title: "unit3_Assignment2_ReadingScores"
author: "Jin Chen"
date: "10/7/2020"
output: html_document
---

```{r}
library(tidyverse)
```
Problem 1.1 - Dataset size

Load the training and testing sets using the read.csv() function, and save them as variables with the names pisaTrain and pisaTest.

How many students are there in the training set? 3,663 

```{r}
pisaTrain <- read_csv("../data/pisa2009train.csv")
pisaTest <- read_csv("../data/pisa2009test.csv")

str(pisaTrain)

```

Problem 1.2 - Summarizing the dataset
Using tapply() on pisaTrain, what is the average reading test score of males? 483.5325
Of females? 512.9406	

```{r}
pisaTrain %>% 
  filter(male == 1) %>% 
  summarize(meanReadingScore = mean(readingScore))

pisaTrain %>% 
  filter(male == 0) %>% 
  summarize(meanReadingScore = mean(readingScore))

```
Problem 1.3 - Locating missing values

Which variables are missing data in at least one observation in the training set?  Everything except grade, male, publicSchool, urban and readingScore

```{r}
library(skimr)
skim(pisaTrain)

```

Problem 1.4 - Removing missing values

Linear regression discards observations with missing data, so we will remove all such observations from the training and testing sets. Later in the course, we will learn about imputation, which deals with missing data by filling in missing values with plausible information.

Type the following commands into your R console to remove observations with any missing value from pisaTrain and pisaTest:
pisaTrain = na.omit(pisaTrain)
pisaTest = na.omit(pisaTest)

How many observations are now in the training set? 2414
How many observations are now in the testing set? 990

```{r}
# pisaTrain = na.omit(pisaTrain)
# pisaTest = na.omit(pisaTest)

pisaTrain <- pisaTrain %>% 
  drop_na()

pisaTest <- pisaTest %>% 
  drop_na()

```

#Problem 2.1 - Factor variables

Factor variables are variables that take on a discrete set of values, like the "Region" variable in the WHO dataset from the second lecture of Unit 1. This is an unordered factor because there isn't any natural ordering between the levels. An ordered factor has a natural ordering between the levels (an example would be the classifications "large," "medium," and "small").

Which of the following variables is an unordered factor with at least 3 levels? (Select all that apply.) raceeeth
Which of the following variables is an ordered factor with at least 3 levels? (Select all that apply.) grade


#Problem 2.2 - Unordered factors in regression models

To include unordered factors in a linear regression model, we define one level as the "reference level" and add a binary variable for each of the remaining levels. In this way, a factor with n levels is replaced by n-1 binary variables. The reference level is typically selected to be the most frequently occurring level in the dataset.

As an example, consider the unordered factor variable "color", with levels "red", "green", and "blue". If "green" were the reference level, then we would add binary variables "colorred" and "colorblue" to a linear regression problem. All red examples would have colorred=1 and colorblue=0. All blue examples would have colorred=0 and colorblue=1. All green examples would have colorred=0 and colorblue=0.

Now, consider the variable "raceeth" in our problem, which has levels "American Indian/Alaska Native", "Asian", "Black", "Hispanic", "More than one race", "Native Hawaiian/Other Pacific Islander", and "White". Because it is the most common in our population, we will select White as the reference level.

Which binary variables will be included in the regression model? (Select all that apply.) Everyone except raceethWhite (which will be the reference variable)

#Problem 2.3 - Example unordered factors

Consider again adding our unordered factor race to the regression model with reference level "White".

For a student who is Asian, which binary variables would be set to 0? All remaining variables will be set to 1. (Select all that apply.) All except raceethAsian
For a student who is white, which binary variables would be set to 0? All remaining variables will be set to 1. (Select all that apply.) All will be set to 0 excep raceethWhite

#Problem 3.1 - Building a model

Because the race variable takes on text values, it was loaded as a factor variable when we read in the dataset with read.csv() -- you can see this when you run str(pisaTrain) or str(pisaTest). However, by default R selects the first level alphabetically ("American Indian/Alaska Native") as the reference level of our factor instead of the most common level ("White"). Set the reference level of the factor by typing the following two lines in your R console:

pisaTrain$raceeth = relevel(pisaTrain$raceeth, "White")

pisaTest$raceeth = relevel(pisaTest$raceeth, "White")

Now, build a linear regression model (call it lmScore) using the training set to predict readingScore using all the remaining variables.

It would be time-consuming to type all the variables, but R provides the shorthand notation "readingScore ~ ." to mean "predict readingScore using all the other variables in the data frame." The period is used to replace listing out all of the independent variables. As an example, if your dependent variable is called "Y", your independent variables are called "X1", "X2", and "X3", and your training data set is called "Train", instead of the regular notation:

LinReg = lm(Y ~ X1 + X2 + X3, data = Train)

You would use the following command to build your model:

LinReg = lm(Y ~ ., data = Train)

What is the Multiple R-squared value of lmScore on the training set? 0.3251


Note that this R-squared is lower than the ones for the models we saw in the lectures and recitation. This does not necessarily imply that the model is of poor quality. More often than not, it simply means that the prediction problem at hand (predicting a student's test score based on demographic and school-related variables) is more difficult than other prediction problems (like predicting a team's number of wins from their runs scored and allowed, or predicting the quality of wine from weather conditions).

```{r}
# pisaTrain$raceeth = relevel(pisaTrain$raceeth, "White")
# pisaTest$raceeth = relevel(pisaTest$raceeth, "White")

pisaTrain <- pisaTrain %>% 
  mutate(raceeth = fct_relevel(raceeth, levels = "White"))

pisaTest <- pisaTest %>% 
  mutate(raceeth = fct_relevel(raceeth, levels = "White"))

pisaReg <- lm(readingScore ~., data = pisaTrain)
summary(pisaReg)

# library(tidymodels)
# library(broom)
# 
# lm_model <-
#     linear_reg() %>%
#     set_mode("regression") %>%
#     set_engine("lm")
# 
# lm_fit1 <- fit(lm_model, readingScore ~ ., data = pisaTrain)
# lm_fit1

```

Problem 3.2 - Computing the root-mean squared error of the model

What is the training-set root-mean squared error (RMSE) of lmScore? 73.36555
```{r}
#recall this from the recitation
# SSE <- sum((PointsPredictions - NBA_test$pts)^2)
# SST <- sum((mean(NBA$pts) - NBA_test$pts)^2)
# RMSE <- sqrt(SSE/nrow(NBA_test))
# RMSE
# glance(lm_fit1$fit) #using tidymodels version. Sigma is RMSE, gets a slighlty different answer but still about 73


sqrt(mean(pisaReg$residuals^2))

```
Problem 3.3 - Comparing predictions for similar students

Consider two students A and B. They have all variable values the same, except that student A is in grade 11 and student B is in grade 9. What is the predicted reading score of student A minus the predicted reading score of student B? 59.09 

Problem 3.4 - Interpreting model coefficients

What is the meaning of the coefficient associated with variable raceethAsian? Predicted difference in the reading score between an Asian student and a white student who is otherwise identical 

Problem 3.5 - Identifying variables lacking statistical significance

Based on the significance codes, which variables are candidates for removal from the model? Select all that apply. (We'll assume that the factor variable raceeth should only be removed if none of its levels are significant.) Preschool, motherHS, motherWork, fatherHS, fatherWork, selfBornUS, motherBornUS, fatherBornUS, englshAtHome,minutesPerWeekEnglish,studentsInEnglish,schoolHasLibrary,urban

Problem 4.1 - Predicting on unseen data

Using the "predict" function and supplying the "newdata" argument, use the lmScore model to predict the reading scores of students in pisaTest. Call this vector of predictions "predTest". Do not change the variables in the model (for example, do not remove variables that we found were not significant in the previous part of this problem). Use the summary function to describe the test set predictions.

What is the range between the maximum and minimum predicted reading score on the test set?

```{r}
predTest <- predict(pisaReg, newdata = pisaTest)
summary(predTest)
637.7 - 353.2


```
Problem 4.2 - Test set SSE and RMSE

What is the sum of squared errors (SSE) of lmScore on the testing set? 5762082
What is the root-mean squared error (RMSE) of lmScore on the testing set? 76.29079

```{r}
# SSE <- sum((PointsPredictions - NBA_test$pts)^2)
# SST <- sum((mean(NBA$pts) - NBA_test$pts)^2)
# RMSE <- sqrt(SSE/nrow(NBA_test))
# RMSE

SSE <- sum((predTest - pisaTest$readingScore)^2)
SSE
SST <- sum((mean(pisaTrain$readingScore) - pisaTest$readingScore)^2)
SST
RMSE <- sqrt(SSE/nrow(pisaTest))
RMSE

```

Problem 4.3 - Baseline prediction and test-set SSE

What is the predicted test score used in the baseline model? Remember to compute this value using the training set and not the test set.
What is the sum of squared errors of the baseline model on the testing set? HINT: We call the sum of squared errors for the baseline model the total sum of squares (SST).

```{r}
baseline <-  mean(pisaTrain$readingScore)
baseline

```

Problem 4.4 - Test-set R-squared

What is the test-set R-squared value of lmScore?

```{r}
R2 <- 1-SSE/SST
R2

```

