---
title: "unit3_Assignment2"
author: "Jin Chen"
date: "10/7/2020"
output: html_document
---

```{r}
library(tidyverse)
```

Problem 1.1 - Creating Our First Model

We are interested in how changes in these variables affect future temperatures, as well as how well these variables explain temperature changes so far. To do this, first read the dataset climate_change.csv into R.

Then, split the data into a training set, consisting of all the observations up to and including 2006, and a testing set consisting of the remaining years (hint: use subset). A training set refers to the data that will be used to build the model (this is the data we give to the lm() function), and a testing set refers to the data we will use to test our predictive ability.

Next, build a linear regression model to predict the dependent variable Temp, using MEI, CO2, CH4, N2O, CFC.11, CFC.12, TSI, and Aerosols as independent variables (Year and Month should NOT be used in the model). Use the training set to build the model.

Enter the model R2 (the "Multiple R-squared" value): 0.7509

```{r}
climate_change <- read_csv("../data/climate_change.csv")
# library(janitor)
# climate_change <- clean_names(climate_change)

climate_change_train <- climate_change %>% 
  filter(Year <= 2006)

climate_change_test <- climate_change %>% 
  filter(Year > 2006)

climate_change_reg <- lm(Temp ~ MEI + CO2 + CH4 + N2O + `CFC-11` + `CFC-12` + TSI + Aerosols, data = climate_change_train)
summary(climate_change_reg)

```

Problem 1.2 - Creating Our First Model

Which variables are significant in the model? We will consider a variable signficant only if the p-value is below 0.05. MEI, CO2, CFC.11, CFC.12, TSI, and Aerosols

Problem 2.1 - Understanding the Model

Current scientific opinion is that nitrous oxide and CFC-11 are greenhouse gases: gases that are able to trap heat from the sun and contribute to the heating of the Earth. However, the regression coefficients of both the N2O and CFC-11 variables are negative, indicating that increasing atmospheric concentrations of either of these two compounds is associated with lower global temperatures.

Which of the following is the simplest correct explanation for this contradiction? All of the gas concentration variables reflect human development - N2O and CFC.11 are correlated with other variables in the data set. 

Problem 2.2 - Understanding the Model

Compute the correlations between all the variables in the training set. Which of the following independent variables is N2O highly correlated with (absolute correlation greater than 0.7)? CO2, CH4, CFC-12

Which of the following independent variables is CFC.11 highly correlated with? CH4, CFC-12

```{r}
cor(climate_change_train)

```

Problem 3 - Simplifying the Model

Given that the correlations are so high, let us focus on the N2O variable and build a model with only MEI, TSI, Aerosols and N2O as independent variables. Remember to use the training set to build the model.

Enter the coefficient of N2O in this reduced model: 0.02532
(How does this compare to the coefficient in the previous model with all of the variables?)

Enter the model R2: 0.7261

```{r}
climate_change_reg2 <- lm(Temp ~ MEI + N2O + TSI + Aerosols, data = climate_change_train)
summary(climate_change_reg2)


```

Problem 4 - Automatically Building the Model

We have many variables in this problem, and as we have seen above, dropping some from the model does not decrease model quality. R provides a function, step, that will automate the procedure of trying different combinations of variables to find a good compromise of model simplicity and R2. This trade-off is formalized by the Akaike information criterion (AIC) - it can be informally thought of as the quality of the model with a penalty for the number of variables in the model.

The step function has one argument - the name of the initial model. It returns a simplified model. Use the step function in R to derive a new model, with the full model as the initial model (HINT: If your initial full model was called "climateLM", you could create a new model with the step function by typing step(climateLM). Be sure to save your new model to a variable name so that you can look at the summary. For more information about the step function, type ?step in your R console.)

Enter the R2 value of the model produced by the step function: 0.7508

```{r}
climate_change_step <- step(climate_change_reg)
summary(climate_change_step)


```

Problem 5 - Testing on Unseen Data

We have developed an understanding of how well we can fit a linear regression to the training data, but does the model quality hold when applied to unseen data?

Using the model produced from the step function, calculate temperature predictions for the testing data set, using the predict function.

Enter the testing set R2: 0.6286051

```{r}
climate_change_prediction <- predict(climate_change_step, newdata = climate_change_test)
summary(climate_change_prediction)

SSE <- sum((climate_change_prediction - climate_change_test$Temp)^2)
SST <- sum((mean(climate_change_train$Temp) - climate_change_test$Temp)^2)
R2 <- 1 - SSE/SST
R2

```

