---
title: "unit2_Assignment2_State"
author: "Jin Chen"
date: "10/8/2020"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
statedata <- read_csv("../data/statedata.csv")
```

# Problem 1.1 - Data Exploration

We begin by exploring the data. Plot all of the states' centers with latitude on the y axis (the "y" variable in our dataset) and longitude on the x axis (the "x" variable in our dataset). The shape of the plot should look like the outline of the United States! Note that Alaska and Hawaii have had their coordinates adjusted to appear just off of the west coast.

In the R command you used to generate this plot, which variable name did you use as the first argument?

```{r}
ggplot(statedata, aes(x = x, y = y)) + 
  geom_point()



```

# Problem 1.2 - Data Exploration

Using the tapply command, determine which region of the US (West, North Central, South, or Northeast) has the highest average high school graduation rate of all the states in the region:
West
```{r}
statedata %>% 
  group_by(state.region) %>% 
  summarize(meansGrad = mean(HS.Grad))
  

```

Problem 1.3 - Data Exploration

Now, make a boxplot of the murder rate by region (for more information about creating boxplots in R, type ?boxplot in your console).

Which region has the highest median murder rate? South 

```{r}
ggplot(statedata, aes(x = state.region, y = Murder)) +
  geom_boxplot()

```

# Problem 1.4 - Data Exploration

You should see that there is an outlier in the Northeast region of the boxplot you just generated. Which state does this correspond to? (Hint: There are many ways to find the answer to this question, but one way is to use the subset command to only look at the Northeast data.) New York 

```{r}
statedata %>% 
  filter(state.region == "Northeast") %>% 
  arrange(desc(Murder))
```

# Problem 2.1 - Predicting Life Expectancy - An Initial Model

We would like to build a model to predict life expectancy by state using the state statistics we have in our dataset.

Build the model with all potential variables included (Population, Income, Illiteracy, Murder, HS.Grad, Frost, and Area). Note that you should use the variable "Area" in your model, NOT the variable "state.area".

What is the coefficient for "Income" in your linear regression model? -0.0000218 

```{r}
library(tidymodels)

lm.model = linear_reg() %>%
  set_engine(engine = "lm") %>%
  set_mode(mode = "regression")

lm.fit <- lm.model %>% 
  fit(Life.Exp ~ Population + Income + Illiteracy + Murder + HS.Grad + Frost + Area, data = statedata)
lm.fit

lifeReg <- lm(Life.Exp ~ Population + Income + Illiteracy + Murder + HS.Grad + Frost + Area, data = statedata)
lifeReg

```

# Problem 2.2 - Predicting Life Expectancy - An Initial Model

Call the coefficient for income x (the answer to Problem 2.1). What is the interpretation of the coefficient x? For a one unit increase in income, predicted life expectancy decreases by |x| correct

# Problem 2.3 - Predicting Life Expectancy - An Initial Model

Now plot a graph of life expectancy vs. income using the command:

plot(statedata$Income, statedata$Life.Exp)

Visually observe the plot. What appears to be the relationship? Life expectancy is somewhat positively correlated with income. correct

```{r}
ggplot(statedata, aes(x = Income, y = Life.Exp)) +
  geom_point()

```

# Problem 2.4 - Predicting Life Expectancy - An Initial Model

The model we built does not display the relationship we saw from the plot of life expectancy vs. income. Which of the following explanations seems the most reasonable? Multicollinearity


# Problem 3.1 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Recall that we discussed the principle of simplicity: that is, a model with fewer variables is preferable to a model with many unnnecessary variables. Experiment with removing independent variables from the original model. Remember to use the significance of the coefficients to decide which variables to remove (remove the one with the largest "p-value" first, or the one with the "t value" closest to zero), and to remove them one at a time (this is called "backwards variable selection"). This is important due to multicollinearity issues - removing one insignificant variable may make another previously insignificant variable become significant.

You should be able to find a good model with only 4 independent variables, instead of the original 7. Which variables does this model contain? Population, Murder, Frost, HS.Grad correct

```{r}
options(scipen = 999) #turn off scientfic notation
options(scipen = 0) #turn on scientific notation

lm.fit.stats = lm.fit %>% 
  tidy(interval = TRUE)
lm.fit.stats
#keep pop, murder, HS.grad, Frost

lm.fit2 <- lm.model %>% 
  fit(Life.Exp ~ Population + Murder + HS.Grad + Frost , data = statedata)
glance(lm.fit2)



```

# Problem 3.2 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Removing insignificant variables changes the Multiple R-squared value of the model. By looking at the summary output for both the initial model (all independent variables) and the simplified model (only 4 independent variables) and using what you learned in class, which of the following correctly explains the change in the Multiple R-squared value?

We expect the "Multiple R-squared" value of the simplified model to be slightly worse than that of the initial model. It can't be better than the "Multiple R-squared" value of the initial model.

Explanation: When we remove insignificant variables, the "Multiple R-squared" will always be worse, but only slightly worse. This is due to the nature of a linear regression model. It is always possible for the regression model to make a coefficient zero, which would be the same as removing the variable from the model. The fact that the coefficient is not zero in the intial model means it must be helping the R-squared value, even if it is only a very small improvement. So when we force the variable to be removed, it will decrease the R-squared a little bit. However, this small decrease is worth it to have a simpler model.

On the contrary, when we remove insignificant variables, the "Adjusted R-squred" will frequently be better. This value accounts for the complexity of the model, and thus tends to increase as insignificant variables are removed, and decrease as insignificant variables are added.

```{r}
glance(lm.fit)
glance(lm.fit2)
```

# Problem 3.3 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Using the simplified 4 variable model that we created, we'll now take a look at how our predictions compare to the actual values.

Take a look at the vector of predictions by using the predict function (since we are just looking at predictions on the training set, you don't need to pass a "newdata" argument to the predict function).

Which state do we predict to have the lowest life expectancy? Alabama 
Which state actually has the lowest life expectancy? South Carolina 

```{r}
lm.predict = lm.fit2 %>% 
  predict(statedata) %>% 
  bind_cols(statedata)
lm.predict

lm.predict %>% 
  arrange(.pred)

lm.predict %>% 
  filter(Life.Exp == min(Life.Exp))

```

# Problem 3.4 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Which state do we predict to have the highest life expectancy? Washington 
Which state actually has the highest life expectancy? Hawaii 

```{r}
lm.predict %>% 
  filter(.pred == max(.pred))

lm.predict %>% 
  filter(Life.Exp == max(Life.Exp))

```

# Problem 3.5 - Predicting Life Expectancy - Refining the Model and Analyzing Predictions

Take a look at the vector of residuals (the difference between the predicted and actual values).

For which state do we make the smallest absolute error? Indiana 
For which state do we make the largest absolute error? Hawaii

```{r}
lm.predict2 <- augment(lm.fit2$fit, data = statedata) %>% 
    rowid_to_column()
lm.predict2 %>% 
  select(rowid, state.name, Life.Exp, .fitted:.std.resid) %>% 
  arrange(.resid)

lm.predict2 %>% 
  select(rowid, state.name, Life.Exp, .fitted:.std.resid) %>% 
  mutate(.residabs = abs(.resid)) %>% 
  arrange(desc(.residabs))

```

