---
title: "Elantra Sales"
author: "Jin Chen"
date: "10/8/2020"
output: html_document
---
```{r}
library(tidyverse)
```


# Problem 1 - Loading the Data

Load the data set. Split the data set into training and testing sets as follows: place all observations for 2012 and earlier in the training set, and all observations for 2013 and 2014 into the testing set.

How many observations are in the training set? 36

```{r}
elantra <- read_csv("../data/elantra.csv")
str(elantra)

elantra_train <- elantra %>% 
  filter(Year <= 2012)

elantra_test <- elantra %>% 
  filter(Year > 2012)

```

# Problem 2.1 - A Linear Regression Model

Build a linear regression model to predict monthly Elantra sales using Unemployment, CPI_all, CPI_energy and Queries as the independent variables. Use all of the training set data to do this.

What is the model R-squared? Note: In this problem, we will always be asking for the "Multiple R-Squared" of the model.

```{r}
library(tidymodels)


```


# Problem 2.2 - Significant Variables
How many variables are significant, or have levels that are significant? Use 0.10 as your p-value cutoff.